MLH1 Results Outline

### 2020-02-15 doc/Results.rmd
de-bugged for knitting, cleaned up outliers for total SC

@@

show_col(hue_pal()(5))

@@

tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

#seems like the outliers have been removed

MLH1.merge.Skel.wsb.male <- MLH1.merge.Skel  %>% filter(category < "WSB male" )

tot.sc.wsb.male <- ggplot(data= MLH1.merge.Skel.wsb.male, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

tot.sc.single.male <- ggplot(data= MLH1.merge.Skel  %>% filter(category == "G female" ) , aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

Checking the outliers at category level,

WSB male,
10apr15_31mar15_WSB_m3_sp1_4_rev.tif

G outliers, 
G 19oct15_12oct15_G_m1_sp1_17.1_rev.tif

LEW, low outlier, 
6mar18_1feb18_LEW_m1_sp1_21.5_rev

PWD outliers, 6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif, 7mar15_1mar15_PWD_m1_sp4_2_rev.tif, and 20jul15_18may15_PWD_m2_sp1_10.2_rev.tif

G female; 1mar16_20feb16_G_f3_sp1_8_rev.tif and 21nov16_16jan16_G_f2_sp1_15_rev.tif

@@

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)

#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)

@@

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(sex~subsp, scales="free")

tot.sc77 <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(0, 800))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

tot.sc_bin <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())  +ylim(c(0, 3000))

#there should be tom

@@

@@

Total SC

```{r show.tot.sc, echo=FALSE}
totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(sex~subsp, scales="free")

tot.sc77 <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(150, 650))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
tot.sc77

tot.sc_bin <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
```

>>>>

# Data set ups

@@

#show_col(hue_pal()(5))

@@

# -- EXAMINE the ggplot scatter plots for outliers
tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")

#round 1
SC.skel.outliers_round1 <- c(
  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",

  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",

       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",

   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
   "6may15_17apr15_PWD_f1_sp1_7_rev",

      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",

    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
 "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
 "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
 "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,

  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",

# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
"16aug19_27aug18_AST_m1_sp1_23_rev.tif",

#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",

#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
 "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
 "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
 "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
 "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
 "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
 "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
 "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )

MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed

remove.mice_round2 <- c(
    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",

  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",

# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",

#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
#SPIC male  x
#SPIC female  x
#TOM x
#AST x
#CZECH male x

#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
 "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
#SPRET female X
#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
)
#3remove.mice_round2

MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]

all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

@@

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
totSC.mouse.av.table <- add_species(totSC.mouse.av.table)

#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)

@@

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av. total SC area")

mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")

@@

```{r Q1.tot.SC_show, echo=FALSE}

Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")

Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
   "WSB","G","LEW","PERC", 
  "PWD", "SKIVE",  "KAZ",
  "MSM", "MOLF"), order=T )

Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])

#order the strains
Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
  arrange(strain) %>%
  mutate(mouse = factor(mouse))

Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])

#add category mean

mouse.av_Tot.SC <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(sex~subsp, scales="free")+ 
  theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(0,3000 ))+ggtitle("Mouse av. total SC area")

mouse.av_Tot.SC
```

<!-- can't figure out the ordering -->

@@

## Total SC

`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images. X observations removed for having extra signal from other cells. `r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)

```{r show.tot.sc, echo=FALSE}

#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, mouse) %>%
#  mutate(file.name = factor(file.name))

mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")
mouse.av_sc_male
